A Compound Decision Approach to Covariance Matrix Estimation

### 2022-03-16 msg.tex
update reference & abstrct

\label{firstpage}
\begin{abstract}
  Covariance matrix estimation is a fundamental statistical task in many applications, but the sample covariance matrix is sub-optimal when the sample size is comparable to or less than the number of features. Such high-dimensional settings are common in modern genomics, where covariance matrix estimation is frequently employed as a method for inferring gene networks. To achieve estimation accuracy in these settings, existing methods typically either assume that the population covariance matrix has some particular structure, for example sparsity, or apply shrinkage to better estimate the population eigenvalues. In this paper, we study a new approach to estimating high-dimensional covariance matrices. We first frame covariance matrix estimation as a compound decision problem. This motivates defining a class of decision rules and using a nonparametric empirical Bayes $g$-modeling approach to estimate the optimal rule in the class. Simulation results and gene network inference in an RNA-seq experiment in mouse show that our approach is comparable to or can outperform a number of state-of-the-art proposals, particularly when the sample eigenvectors are poor estimates of the population eigenvectors.
\end{abstract}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textheight]{img/networks_A_2.png}
  \end{center}
\caption{Amygdala gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge.}
\label{network_A}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textheight]{img/networks_FC_2.png}
  \end{center}
\caption{Frontal cortex gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge.}
\label{network_FC}
\end{figure}

>>>>

\label{firstpage}
\begin{abstract}
  Covariance matrix estimation is a fundamental statistical task in many applications, but the sample covariance matrix is sub-optimal when the sample size is comparable to or less than the number of features. Such high-dimensional settings are common in modern genomics, where covariance matrix estimation is frequently employed as a method for inferring gene networks. To achieve estimation accuracy in these settings, existing methods typically either assume that the population covariance matrix has some particular structure, for example sparsity, or apply shrinkage to better estimate the population eigenvalues. In this paper, we study a new approach to estimating high-dimensional covariance matrices. We first frame covariance matrix estimation as a compound decision problem. This motivates defining a class of decision rules and using a nonparametric empirical Bayes $g$-modeling approach to estimate the optimal rule in the class. Simulation results and gene network inference in an RNA-seq experiment in mouse show that our approach is comparable to or can outperform a number of state-of-the-art proposals.
\end{abstract}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textheight]{img/networks_A_2.png}
  \end{center}
\caption{Amygdala gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge. This figure appears in color in the electronic version of this article, and any mention of color refers to that version.}
\label{network_A}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=0.45\textheight]{img/networks_FC_2.png}
  \end{center}
\caption{Frontal cortex gene networks recovered by the top performers in Table \ref{tab:tab1}. Line types indicate the method(s) that recovered each each edge. This figure appears in color in the electronic version of this article, and any mention of color refers to that version.}
\label{network_FC}
\end{figure}
